---
title: "Reproducible Research: Peer Assessment 1"
author: "Maarten Caminada"
date: "Tuesday, November 11, 2014"
output: 
        html_document:
                keep_md: true
---

## Loading and preprocessing the data

First we need to load the data, clean up the dataframe for NAs and load various libraries.
```{r}
wd <- "c:/maca/Rdata/RepRes1" #the directory where I unzipped the datafiles
setwd(wd) #is now my working directory

Sys.setlocale("LC_TIME", "C") # change timesettings to English, otherwise the names of the days appeared in Dutch

library(knitr) #load knitr, assuming the package is installed
library(markdown) # dito for markdown
library(plyr)
library(lubridate)
library(ggplot2)

activity <- read.csv("activity.csv") 
act_cc <- activity[complete.cases(activity),] #only use the rows without missing values

```

## What is mean total number of steps taken per day?

Make the histogram and calculate the mean and median.

```{r}
totals <- ddply(act_cc, .(date), summarize, steps=sum(steps))
hist(totals$steps)
summary(act_cc) 
        
```


## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
